A Simplified Plackett-Luce Likelihood Function for Rank Orders

### 2024-09-18 inst/hankin-AustrianJournalofStatistics.tex
typos

\noindent where $B=1/b$.  Note carefully that these expressions are
likelihoods, not probabilities, and do not sum to 1.  Previous related
work \citep{hankin2024_hyper3} presented the case $n=3$ [in current
  notation].  Considering the case $n=8$ gives us $8+1=9$ distinct
likelihood functions, Figure~\ref{ninelikes}.  We see an unanticipated
asymmetry between $r$ and $n-r$.  For $r=5$ one might expect that
$\hat{a}=\frac{1}{2}$ on the grounds that the focal competitor's
performance was median.  However, $\hat{a}\simeq 0.5886$.
Figure~\ref{dotprobs} shows the maximum likelihood estimator $\hat{a}$
for various values of $n,r$, obtained by direct numerical maximization
of equation~\ref{likeforrn3}.

@@

Is there evidence that the author's performance differs between
years?  Figure~\ref{parkruncontour} shows contours of
support [again normalised so the support of the evaluate is zero] for
the parkrun dataset, as a function of 2023 strength on the
horizontal axis and 2024 strength on the vertical axis.  A null of
equal strengths appears as the diagonal line.  We see that the maximum
likelihood estimate is $\hat{p}_\mathrm{2023}=0.427$,
$\hat{p}_\mathrm{2024}=0.408$.  However, a constrained optimization
shows that the maximum support along the null diagonal is about
$-2.43$, at ${p}_\mathrm{2023}={p}_\mathrm{2024}\simeq 0.409$.  The
two units of support criterion~\citep{edwards1972} is met, and we are
justified in concluding that the student is indeed more competent at
applied than pure mathematics.  Alternatively, we may observe that
$2\mathcal{S}=4.86$ is in the tail region of its asymptotic $\chi^2_1$
distribution, corresponding to a $p$-value of about $0.028$.

Now consider table~\ref{fisherparkrun}, giving a summary of whether
nonfocal ocmpetitors place ahead of the author, split by venue.  A
two-sided Fisher test gives a $p$-value of about $5\times 10^{-14}$,
surely an unreasonable conclusion given the author's perceived
performance stagnation.

>>>>

\noindent where $B=1/b$.  Note carefully that these expressions are
likelihoods, not probabilities, and do not sum to 1.  Previous related
work \citep{hankin2024_hyper3} presented the case $n=3$ [in current
  notation] but here arbitrary $n$ may be considered.  Taking the case
$n=8$, for example, gives us $8+1=9$ distinct likelihood functions,
Figure~\ref{ninelikes}.  We see an unanticipated asymmetry between $r$
and $n-r$.  For $r=5$ one might expect that $\hat{a}=\frac{1}{2}$ on
the grounds that the focal competitor's performance was median.
However, $\hat{a}\simeq 0.5886$.  Figure~\ref{dotprobs} shows the
maximum likelihood estimator $\hat{a}$ for various values of $n,r$,
obtained by direct numerical maximization of
equation~\ref{likeforrn3}.

@@

Is there evidence that the author's performance differs between years?
Figure~\ref{parkruncontour} shows contours of support [again
  normalised so the support of the evaluate is zero] for the parkrun
dataset, as a function of 2023 strength on the horizontal axis and
2024 strength on the vertical axis.  A null of equal strengths appears
as the diagonal line.  We see that the maximum likelihood estimate is
$\hat{p}_\mathrm{2023}=0.413$, $\hat{p}_\mathrm{2024}=0.485$.
However, a constrained optimization shows that the maximum support
along the null diagonal is about $-0.22$, at
${p}_\mathrm{2023}={p}_\mathrm{2024}\simeq 0.441$.  The two units of
support criterion~\citep{edwards1972} is not met, and we conclude that
there is no evidence for a change in competitive strength between the
years.  Now consider table~\ref{fisherparkrun}, giving a summary of
whether nonfocal competitors place ahead of the author, split by year.
A two-sided Fisher test gives a $p$-value of about $5\times 10^{-14}$,
surely an unreasonable conclusion given the author's perceived
performance stagnation.
